### 2020-02-18 thesis/content/chapter-related-work.tex
continued wl section

The \acf{wl} algorithm~\cite{Weisfeiler1968} characterizes a graph $G = (\mathcal{V}, \mathcal{E})$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
A mapping $\chi_{G, k}: \mathcal{V}^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G = (\mathcal{V}, \mathcal{E})$.
We say  $\chi'$ \textit{refines} $\chi$ ($\chi' \succeq \chi$) iff.\ $\forall a, b \in \mathcal{V}^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.

@@

\subsubsection{Discriminative Power of \acs{wl}}
Now we will look at the types of graphs that can be distinguished by \acs{wl} in relation to the \acs{wl}-dimension $k$.
\begin{lem}
	$G \mathrel{\not\simeq_k} H \implies G \mathrel{\not\simeq_{k+1}} H$ (the \acs{wl} power is never lower for higher $k$).
\end{lem}
\begin{hproof}
	All $k$-tuples can be mapped to $(k+1)$-tuples via $\varphi(v_1, \dots, v_k) \coloneqq (v_1, \dots,\allowbreak v_k, v_k)$.
	For each neighbor $(v[u/1], \dots,\allowbreak v[u/k])$ of $v$, there is a corresponding neighbor $(\varphi(v[u/1]), \dots \varphi(v[u/k]), \varphi(v[u/k]))$ of $\varphi(v)$.
	Using \cref{eq:related:wlk-init} (for $i = 0$) and \cref{defn:related:wlk-refine} (for $i > 0$) it follows that $\forall v, w \in \mathcal{V}_G^k: \chi_{G, k}^i(v) \neq \chi_{G, k}^i(w) \rightarrow \chi_{G, k+1}^i(\varphi(v)) \neq \chi_{G, k+1}^i(\varphi(w))$.
	The lemma then follows by \cref{defn:related:wl-distinguishable}.
\end{hproof}
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/related-work/wl2-problem-solution.pdf}
	\caption{
		Two non-isomorphic graphs with $G \mathrel{\simeq_1} H$ and $G \mathrel{\not\simeq_2} H$.
	}\label{fig:related:wl2-problem-solution}
\end{figure}
\Cref{fig:related:wl2-problem-solution} illustrates the higher discriminative power of higher \acs{wl}-dimensions;
it shows how 2-WL is able to distinguish the two 1-WL indistinguishable graphs from \cref{fig:related:wl1-problem}.

@@

\paragraph{Graph Embeddings}
Skip-gram methods can not only be used for vertex embeddings but also to embed entire graphs.
One way to do this is via the \texttt{graph2vec}~\cite{Narayanan2017} method.
It is inspired by \texttt{doc2vec}~\cite{Le2014} which in turn is based on \texttt{word2vec}.
\texttt{graph2vec} gets a set of graphs $\mathbb{G} = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $f(G_1), \dots, f(G_N)$.
While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathbb{G}$ as context roots.
The context $C_k(G_i)$ of a graph $G_i = (\mathcal{V}_i, \mathcal{E}_i)$ is defined as
\begin{equation}
	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_i}}\, {\left\{ {\textrm{WL}_l(G_i)}_j \right\}}_{l = 0}^k % chktex 21
	\text{ with } \textrm{WL}_l(G) \coloneqq \text{$l$ iter.\ \acs{wl}-1 coloring of $G$.}
\end{equation}
Intuitively this context can be understood as the set of \acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2k$.
Since \ac{wl} is used to identify distinct subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
Using the previous definitions, the context root embedding function has the signature $f: \mathbb{G} \to \mathbb{R}^d$ while the context element embedding function is of type $f_C: \bigcup_{G_i \in \mathbb{G}} C_k(G_i) \to \mathbb{R}^d$.
To find those embeddings the \texttt{word2vec} objective from \cref{eq:related:gcr:embed:word2vec} is reused.
Analogous to \texttt{word2vec}, \texttt{graph2vec} therefore embeds graphs that share subgraphs close to each other, whereas graphs that do not share substructures tend to be embedded further away from each other.

>>>>

The \acf{wl} algorithm~\cite{Weisfeiler1968}\cite{Cai1992} characterizes a graph $G = (\mathcal{V}, \mathcal{E})$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
A mapping $\chi_{G, k}: \mathcal{V}^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G = (\mathcal{V}, \mathcal{E})$.
We say  $\chi'$ \textit{refines} $\chi$ ($\chi' \succeq \chi$) iff.\ $\forall a, b \in \mathcal{V}^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.

@@

\subsubsection{Discriminative Power of \acs{wl}}
Now we will look at the types of graphs that can be distinguished by \acs{wl} in relation to the \acs{wl}-dimension $k$.
\begin{lem}\label{lem:related:wl-lower}
	$G \mathrel{\not\simeq_k} H \implies G \mathrel{\not\simeq_{k+1}} H$ ($(k+1)$-\acs{wl} is at least as powerful as $k$-\acs{wl}).
\end{lem}
\begin{hproof}
	All $k$-tuples can be mapped to $(k+1)$-tuples via $\varphi(v_1, \dots, v_k) \coloneqq (v_1, \dots,\allowbreak v_k, v_k)$.
	For each neighbor $(v[u/1], \dots,\allowbreak v[u/k])$ of $v$, there is a corresponding neighbor $(\varphi(v[u/1]), \dots \varphi(v[u/k]), \varphi(v[u/k]))$ of $\varphi(v)$.
	Using \cref{eq:related:wlk-init} (for $i = 0$) and \cref{defn:related:wlk-refine} (for $i > 0$) it follows that $\forall v, w \in \mathcal{V}_G^k: \chi_{G, k}^i(v) \neq \chi_{G, k}^i(w) \rightarrow \chi_{G, k+1}^i(\varphi(v)) \neq \chi_{G, k+1}^i(\varphi(w))$.
	The lemma then follows by \cref{defn:related:wl-distinguishable}.
\end{hproof}
\begin{fact}[see~\citet{Immerman1990}]\label{fact:related:wl-upper}
	For all $k$ there are non-isomorphic graphs $G \not\simeq H$ of size $\mathcal{O}(k)$ with $G \mathrel{\simeq_k} H$.
\end{fact}
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/related-work/wl2-problem-solution.pdf}
	\caption{
		Two non-isomorphic graphs with $G \mathrel{\simeq_1} H$ and $G \mathrel{\not\simeq_2} H$.
	}\label{fig:related:wl2-problem-solution}
\end{figure}
\begin{cor}\label{cor:related:wl-monotonous}
	The discriminative power of $k$-\acs{wl} grows monotonously with $k$ and never converges, i.e.\ for all $k \in \mathbb{N}$ there is an $l \in \mathbb{N}$ s.t.\ the set of $k$-\ac{wl} distinguishable graphs is a true subset of the $(k+l)$-\ac{wl} distinguishable graphs.
\end{cor}
\begin{hproof}
	Note that $k$-\acs{wl} trivially solves the graph isomorphism problem for all graphs of size $\leq k$ via the initial coloring (see~\cref{eq:related:wlk-init}).
	Using \cref{lem:related:wl-lower} and \cref{fact:related:wl-upper} the corollary directly follows.
\end{hproof}
\Cref{fig:related:wl2-problem-solution} illustrates \cref{cor:related:wl-monotonous} by showing how 2-WL is able to distinguish the two 1-WL indistinguishable graphs from \cref{fig:related:wl1-problem}.
Since the time complexity of $k$-\acs{wl} grows exponentially with $k$~\cite[corr.~1.9.7]{Immerman1990}, a trade-off between the runtime and the discriminative power has to be made.

@@

\paragraph{Graph Embeddings}
Skip-gram methods can not only be used for vertex embeddings but also to embed entire graphs.
One way to do this is via the \texttt{graph2vec}~\cite{Narayanan2017} method.
It is inspired by \texttt{doc2vec}~\cite{Le2014} which in turn is based on \texttt{word2vec}.
\texttt{graph2vec} gets a set of graphs $\mathbb{G} = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $f(G_1), \dots, f(G_N)$.
While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathbb{G}$ as context roots.
The context $C_k(G_i)$ of a graph $G_i = (\mathcal{V}_i, \mathcal{E}_i)$ is defined as
\begin{equation}
	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_i}}\, {\left\{ {\chi_{G_i, 1}^l(v_j) \right\}}_{l = 0}^k % chktex 21
	\text{ with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^l$ as in \cref{defn:related:wl1-refine}.}
\end{equation}
Intuitively this context can be understood as the set of 1-\acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2k$.
Since \ac{wl} is used to identify distinct subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
Using the previous definitions, the context root embedding function has the signature $f: \mathbb{G} \to \mathbb{R}^d$ while the context element embedding function is of type $f_C: \bigcup_{G_i \in \mathbb{G}} C_k(G_i) \to \mathbb{R}^d$.
To find those embeddings the \texttt{word2vec} objective from \cref{eq:related:gcr:embed:word2vec} is reused.
Analogous to \texttt{word2vec}, \texttt{graph2vec} therefore embeds graphs that share subgraphs close to each other, whereas graphs that do not share substructures tend to be embedded further away from each other.
